Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
minor modifications on paper

Another obstacle to trying to parallelize numerical methods is that the methods
are inherently serial in time---in that evaluation of the next time point $n+1$
depends on the previous values, say those at time $n-1$ and $n$. This setup
does not allow for parallelism. The only way to incorporate parallelism is to
create schemes that have parallelizable components---such as an update or a
refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
methodology section, one such scheme would be the parareal algorithm. As an
overview, the parareal algorithm first does a coarse (low-order and fast)
approximation, then seeks to iteratively correct using smaller, more refined
numerical methods done in parallel from approximated starting points
approximated by the coarse solution.

@@

\begin{equation}
\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \times T(n)} \times 100\%
\end{equation}

@@

\begin{equation}
\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \times 100\%
\end{equation}

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$

@@

We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
Our coarse operator was forward Euler step, with 100 steps, while the fine
operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
factor. We tested the Parareal algorithm on two sets of differential equations:

>>>>

Another obstacle to trying to parallelize numerical methods is that the methods
are inherently serial in time---in that evaluation of the next time point $n+1$
depends on the previous values, say those at time $n-1$ and $n$. This setup
does not allow for parallelism. The only way to incorporate parallelism is to
create schemes that have parallelizable components---such as an update or a
refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
metodology section, one such scheme would be the parareal algorithm. As an
overview, the parareal algorithm first does a coarse (low-order and fast)
approximation, then seeks to iteratively correct using smaller, more refined
numerical methods done in parallel from approximated starting points
approximated by the coarse solution.

@@

\begin{equation}
\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \cdot T(n)} \cdot 100\%
\end{equation}

@@

\begin{equation}
\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \cdot 100\%
\end{equation}

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$

@@

We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
Our coarse operator was forward Euler step, with 100 steps, while the fine
operator was forward Euler, with $Q \cdot 100$ steps, where $Q$ is the quality
factor. We tested the Parareal algorithm on two sets of differential equations:
